---
title: "Time Series"
author: "Alain Kuiete"
date: "6/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the data
```{r}
tute1 <- read.csv("tute1.csv", header=TRUE)
View(tute1)
```

Convert data to time series
```{r}
mytimeseries <- ts(tute1[,-1], start=1981, frequency=4)
```


Construct a time series plots of each of the three series
```{r}
autoplot(mytimeseries, facets=TRUE)
```

Load the dataset
```{r}
retaildata <- readxl::read_excel("retail.xlsx", skip=1)
```

Turnover ;  New South Wales ;  Electrical and electronic goods retailing ;
```{r}
myts <- ts(retaildata[,"A3349336V"],
  frequency=12, start=c(1982,4))
```

```{r}
str(myts)
```


autoplot(), ggseasonplot(), ggsubseriesplot(), gglagplot(), ggAcf()
```{r}
autoplot(myts)
```

```{r}
ggseasonplot(myts, polar = TRUE)
```

The polar coordinate shows that there is a large jump on sales in December. The month of June presents also a smooth increase in sales. Conversely, there are fall on sales in Februry and August.

```{r}
ggsubseriesplot(myts)
```


```{r}
gglagplot(myts)
```

```{r}
ggAcf(myts)
```

The scalloped shape is due to seasonality
There are two type of trends: 
The trend of increasing amplitude of seasonalty 















